The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-07-03 Paper/jaccardbaetica.tex
add new changes xavi

Advances in the research of the Roman studies have currently led an environment with more diverse interpretations on commercial dynamics \citep{duncan1982economy,
temin_economy_2006,
quantifyingwilson2009}\xavi{more diverse que qué?}. The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the complexity of the Roman production \xavi{en qué sentido? qué quieres decir aquí con complexity? Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.\citep{brughmans_roman_2016,
orengo_seeds_2016,bayesian_2018,
coto-sarmiento_identifying_2018,
rubio-campillo_ecology_2018}.

@@

This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps. Specifically, our work pretends to detect microeconomic processes\xavi{no me gusta lo de microeconomic, porque es un campo específico de la economía que nada tiene que ver con analizar un limes entero. Quizás regional?} focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

@@

The chronology of the workshops is widely diverse from the first to the third centuries AD\xavi{qué quiere decir que la cronología es diversa?} \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?} \citep{remesal_anforas_2004}.

@@

These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion\xavi{qué information? El código de letras?}. Stamps often displayed a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

\section{Case study: consumption centres}

\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}

The creation of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes\xavi{no entiendo esto...allowed o required?}. This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex. Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply. The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}\xavi{pero estabas hablando del ejército no? Has hecho un salto aquí a los ciudadanos}.

@@

Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}. The high percentage of amphorae\xavi{amphorae o Dressel 20?} found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.

@@

The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps). Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces. This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border. Specifically, for the German \textit{liomes} we included sites from the two provinces (Ulterior and Citerior\xavi{no es superior e inferior?}) while the analysis of Britannia extended to \textit{Caledonia}.

@@

\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}

@@

This work hasd explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas. For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres. The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.

@@

Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps. In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}. In \textit{Britannia} the majority of similarity stamps were mainly found in military centres. This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres\xavi{pero no decías que el aceite llegaba por las legione? Entonces por qué es un bias? En todo caso ponlo al final de la sección y no al inicio}.

@@

\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution. Nor it was detected consumption areas where stamps are specified from an amphora workshop.

@@

A large number of Dressel 20 were not marked with any stamps, so some researchers suggest that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}. This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}\xavi{a branch? que es una branch?}. This organisation\xavi{this organisation o unique stamp codes?} could have also served to identify different groups of potters working in the same amphora workshop. Potters could have marked the amphorae to distinguish different groups working at the same time in different orders~\citep{li_crossbows_2014}. This hypothesis could explain why we detect different stamps in the same workshop. In any case, at present we do not have enough archaeological evidence to validate which of these interpretations is more plausible in our case study.

>>>>

Advances in the research of the Roman studies have currently led an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
temin_economy_2006,
quantifyingwilson2009}.
%\xavi{more diverse que qué?} 
The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of the Roman production %\xavi{en qué sentido? qué quieres decir aquí con complexity? Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.
\citep{brughmans_roman_2016,
orengo_seeds_2016,bayesian_2018,
coto-sarmiento_identifying_2018,
rubio-campillo_ecology_2018}.

@@

This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps. Specifically, our work pretends to detect regional processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

@@

The chronology of the workshops is wide spanning from the first to the third centuries AD
%\xavi{qué quiere decir que la cronología es diversa?} 
\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?} \citep{remesal_anforas_2004}.

@@

These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion
%\xavi{qué information? El código de letras?}. 
Stamps often displayed a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

\section{Consumption centres: Britannia and Germania}

%\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}

The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
%\xavi{no entiendo esto...allowed o required?}. 
This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex. Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply. The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}\xavi{pero estabas hablando del ejército no? Has hecho un salto aquí a los ciudadanos}.

@@

Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}. The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.

@@

The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps). Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces. This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border. Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of Britannia extended to \textit{Caledonia}.

@@

%\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}

@@

This work has explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas. For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres. The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.

@@

Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps. In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}. In \textit{Britannia} the majority of similarity stamps were mainly found in military centres. This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres\xavi{pero no decías que el aceite llegaba por las legione? Entonces por qué es un bias? En todo caso ponlo al final de la sección y no al inicio}.

@@

%\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution. Nor it was detected consumption areas where stamps are specified from an amphora workshop.

@@

A large number of Dressel 20 were not marked with any stamps, so some researchers suggest that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}. This method would be used as an identifier to count the number of amphorae of a batch \citep{juanmorostesis}.

Unique stamp codes could have also served to identify different groups of potters working in the same amphora workshop. Potters could have marked the amphorae to distinguish different groups working at the same time in different orders~\citep{li_crossbows_2014}. This hypothesis could explain why we detect different stamps in the same workshop. In any case, at present we do not have enough archaeological evidence to validate which of these interpretations is more plausible in our case study.
